This is a team assignment; each team should complete and turn in a PDF created from an Rmd via Github. Please include code and output for the following components:

Description of dataset
What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
The data source is from the California Smokers Cohort (CSC) 2011 through UC San Diego. These data are from a survey designed to investigate factors associated with tobacco quitting behaviors, sponsored/funded by CDPH.

How does the dataset relate to the group problem statement and question?
Through these data, CDPH will better understand tobacco use and behaviors among smokers in California and be able to design an implementation strategies in high-risk communities to increase quitting behaviors.

Import statement
NOTE: Please use datasets available in the PHW251 Project Data github repo (Links to an external site.) (this is important to make sure everyone is using the same datasets)
Use appropriate import function and package based on the type of file


```{r}

library(tidyverse)
library(readr)

```

```{r}

#creating a file path
file_path <- "https://raw.githubusercontent.com/PHW290/phw251_projectdata/main/ca_csc_smoker_data.csv"


#importing full file to see what it looks like
smoker_data <- read_csv(file_path)

```


Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
Document the import process
```{r}

```


Identify data types for 5+ data elements/columns/variables

Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.

Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
```{r}

```


Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type? 
```{r}

```


Provide a basic description of the 5+ data elements
Numeric: mean, median, range
Character:  unique values/categories
Or any other descriptives that will be useful to the analysis
```{r}

```

