**Milestone #3, due Nov 7**
**Clean dataset with descriptive statistics for relevant data elements**
**Please turn in a PDF created from an Rmd with the following components**

**Subset rows or columns, as needed**
```{r}

```

**Create new variables needed for analysis (minimum 2)**
**New variables should be created based on existing columns; for example**
```{r}

```

**Calculating a rate**
```{r}

```

**Combining character strings**
```{r}

```

**If no new values are needed for final tables/graphs, please create 2 new variables anyway**
```{r}

```

**Clean variables needed for analysis (minimum 2)**
Examples
Recode invalid values
Handle missing fields
Recode categories
```{r}

```

**If not needed for final analysis, please create at least 2 new variables anyway**
```{r}

```

**Data dictionary based on clean dataset (minimum 4 data elements), including:**
Variable name
Data type
Description
One or more tables with descriptive statistics for 4 data element
```{r}

```

**PDF that is professionally prepared for presentation**

**Each part of the milestone is clearly on one page (use \newpage to push to a new page)**

**Only the necessary information is outputted (you should suppress, for example, entire data frame outputs)**

**Use of headers and sub headers to create an organized document**